### 2011-11-24 AffineLieAlgebras/Thesis/1.tex
Thesis partially structure cleaned, some translations in chapter 2.

\chapter{Аффинные алгебры Ли}
\label{cha:affine-lie-algebras}
Аффинные алгебры Ли - это естественное обобщение простых конечномерных алгебр Ли. Они возникают при изучении различных физических моделей, например, моделей Весса-Зумино-Новикова-Виттена в конформной теории поля и интегрируемых спиновых цепочек. 
\section{Введение. Простые алгебры Ли.}

@@

Now consider an algebra $\gf$ and a reductive subalgebra
$\af\subset \gf$. Simple roots $\beta_{i}$ of the subalgebra $\af$
can be presented as linear combinations of $\gf$-algebra roots
$\alpha_{j}$: $\beta_{i}=\sum_{j=1,\dots,r_{\gf}}k_{j}
\alpha_{j},\ j=1,\dots,r_{\af}$.

Each irreducible $\gf$-module is also an $\af$-module, although
$L^{\mu}_{\gf}$ is in general not irreducible as $\af$-module. 
It can be
decomposed into a direct sum of irreducible $\af$-modules:
\begin{equation}
  \label{eq:16}
  L^{\mu}_{\gf}=\bigoplus_{\nu}b^{\mu}_{\nu}L^{\nu}_{\af}
\end{equation}
Coefficients in such a decomposition are called branching
coefficients.

It is possible to calculate branching coefficients by constructing and successively  subtracting the submodules $L^{\nu}_{\af}$. 
This traditional approach has serious limitations especially in case of affine Lie algebras. We discuss them in the end of section \ref{sec:comp-algor}.

Now we describe an alternative approach which is based on recurrent properties of branching coefficients. 
But before we proceed to these recurrent relations we need several additional definitions.

For a subalgebra $\af\subset \gf$ we introduce the subalgebra
$\afb$. Consider the root subspace $\hf_{\perp \af}^{\ast }$
orthogonal to $\hf_{\af}$,
\begin{equation*}
\hf_{\perp \af}^{\ast }:=\left\{ \eta \in \hf^{\ast }|\forall
h\in \hf_{\af};\eta \left( h\right) =0\right\} ,
\end{equation*}
and the roots (correspondingly -- positive roots) of $\gf$ orthogonal
to roots of $\af$,
\begin{eqnarray}
\Delta _{\afb} &:&=\left\{ \beta \in \Delta _{\gf}|\forall
h\in \hf_{\af};\beta \left( h\right) =0\right\} ,
\label{delta a ort} \\
\Delta _{\afb}^{+} &:&=\left\{ \beta ^{+}\in \Delta _{\gf%
}^{+}|\forall h\in \hf_{\af};\beta ^{+}\left( h\right) =0\right\} .
\notag
\end{eqnarray}
Let $W_{\afb}$ be a subgroup of $W$ generated by
reflections $w_{\beta }$ with the roots $\beta \in \Delta _{\af_{\perp
}}^{+}$. The subsystem $\Delta _{\afb}$ determines a 
subalgebra $\afb$ with the Cartan subalgebra $\hf_{\af%
_{\perp }}$.

The Cartan subalgebra $\frak{h}$ can be decomposed in the following way:  $\frak{h}=\frak{\frak{h}_{\af}}\oplus
\frak{h}_{\afb}\oplus \frak{h}_{\perp }$

We also introduce the notations
\begin{eqnarray}
\widetilde{\frak{a}_{\perp }} :=\frak{a}_{\perp }\oplus \frak{h}_{\perp }
\qquad
\widetilde{\frak{a}} :=\frak{a}\oplus \frak{h}_{\perp }.
\end{eqnarray}

For $\af$ and $\afb$ we consider the
corresponding Weyl vectors, $\rho _{\af}$ and $\rho _{\af_{\perp
}} $ and compose the so called ''defects'' $\mathcal{D}_{\af}$ and $\mathcal{%
D}_{\afb}$ of the injection:
\begin{equation}
\mathcal{D}_{\af}:=\rho _{\af}-\pi _{\af}\rho , \qquad
\mathcal{D}_{\afb}:=\rho _{\afb}-\pi _{\af%
_{\perp }}\rho .  \label{defect-ort}
\end{equation}

%%!!! Define P^+
For $\mu \in P^{+}$ consider the linked weights $\left\{ \left(
w(\mu +\rho )-\rho \right) |w\in W\right\} $ and their projections
to
$h_{\afb}^{\ast }$ additionally shifted by the defect $-%
\mathcal{D}_{\afb}$:
\begin{equation*}
\mu _{\afb}\left( w\right) :=\pi _{\afb}\left[
w(\mu +\rho )-\rho \right] -\mathcal{D}_{\afb},\quad w\in W.
\end{equation*}
Among the weights $\left\{ \mu _{\af_{\perp
}}\left( w\right) |w\in W\right\} $ one can always choose those located in
the fundamental chamber $\overline{C_{\afb}}$. Let $U$ be the
set of representatives $u$ for the classes $W/W_{\afb}$ such
that

\begin{equation}
U:=\left\{ u\in W|\quad \mu _{\afb}\left( u\right) \in
\overline{C_{\afb}}\right\} \quad .  \label{U-def}
\end{equation}
Thus we can form the subsets:
\begin{equation}
\mu _{\widetilde{\mathfrak{a}}}\left( u\right) :=\pi _{\widetilde{%
\mathfrak{a}}}\left[ u(\mu +\rho )-\rho \right] +\mathcal{D}_{\af%
_{\perp }},\quad u\in U,  \label{mu-a}
\end{equation}
and
\begin{equation}
\mu _{\afb}\left( u\right) :=\pi _{\afb}\left[
u(\mu +\rho )-\rho \right] -\mathcal{D}_{\afb},\quad u\in U.
\label{mu-a-tilda}
\end{equation}

Notice that the subalgebra $\mathfrak{a}_{\bot}$ is regular by definition
since it is built on a subset of roots of the algebra $\mathfrak{g}$.

Denote by  $k_{\xi }^{\left( \mu \right) }$ signed branching coefficients. If $\xi\in \bar C_{\af}$ is in main Weyl chamber $k_{\xi}^{(\mu)}=b^{(\mu)}_{\xi}$ otherwise $k_{\xi}^{(\mu)}=\epsilon(w) b^{(\mu)}_{w (\xi+\rho_{\af})-\rho_{\af}}$ where $w\in W_{\af}$ is such that $w (\xi+\rho_{\af})-\rho_{\af}\in \bar C_{\af}$.

Now we can use the Weyl character formula to write a
recurrent relation \cite{2010arXiv1007.0318L} for signed branching
coefficients $k_{\xi }^{\left( \mu \right) }$ corresponding to an
injection $\af\hookrightarrow \gf$:
\begin{equation}
\begin{array}{c}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{u\in U}\epsilon (u)\;\dim \left( L_{\afb}^{\mu _{\af%
_{\perp }}\left( u\right) }\right) \delta _{\xi -\gamma _{0},\pi _{%
\widetilde{\af}}(u(\mu +\rho )-\rho )}+\right.  \\
\left. +\sum_{\gamma \in \Gamma _{\widetilde{\af}\rightarrow \gf%
}}s\left( \gamma +\gamma _{0}\right) k_{\xi +\gamma }^{\left( \mu \right)
}\right) .
\end{array}
\label{recurrent-rel}
\end{equation}
The recursion is governed by the set $\Gamma _{\af\rightarrow \gf}$ called the injection fan. The latter is defined by the
carrier set $\left\{ \xi \right\} _{\af\rightarrow \gf}$ for the
coefficient function $s(\xi )$
\begin{equation*}
\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}:=\left\{
\xi \in P_{\widetilde{\af}}|s(\xi )\neq 0\right\}
\end{equation*}
appearing in the expansion
\begin{equation}
\prod_{\alpha \in \Delta ^{+}\setminus \Delta _{\bot }^{+}}\left( 1-e^{-\pi
_{\widetilde{\af}}\alpha }\right) ^{\mathrm{mult}(\alpha )-\mathrm{mult}%
_{\af}(\pi _{\widetilde{\af}}\alpha )}=-\sum_{\gamma \in P_{%
\widetilde{\af}}}s(\gamma )e^{-\gamma };\quad
\end{equation}
The weights in $\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}$ are to be shifted by $\gamma _{0}$ -- the lowest vector in $\left\{ \xi
\right\} $ -- and the zero element is to be eliminated:
\begin{equation}
\Gamma _{\af\rightarrow \gf}=\left\{ \xi -\gamma
_{0}|\xi \in \left\{ \xi \right\} \right\} \setminus \left\{ 0\right\} .
\end{equation}
Formula (\ref{eq:14}) is a particular case of recurrent relation for branching coefficients (\ref{recurrent-rel}) in the case of Cartan subalgebra $\af=\hfg$.

If the root system of $\afb$ is generated by some subset of $\gf$
simple roots $\alpha_{1},\dots,\alpha_{r}$ then the recurrent
relation (\ref{recurrent-rel}) is connected with the generalized
Bernstein-Bernstein-Gelfand resolution for parabolic Verma modules
\cite{2011arXiv1102.1702L}.

Another particular case of this formula is connected with tensor
product decompositions. Consider the tensor product of two
irreducible $\gf$-modules $L^{\mu}\otimes L^{\nu}$. It is also a
$\gf$-module but not irreducible in general. So
\begin{equation}
  \label{eq:19}
  L^{\mu}\otimes L^{\nu}=\bigoplus_{\gamma} f^{\mu\nu}_{\gamma}L^{\gamma}
\end{equation}
The coefficients $f^{\mu\nu}_{\gamma}$ are called fusion
coefficients. The problem of computation of fusion coefficients is
equivalent to branching problem for the diagonal subalgebra
$\gf\subset \gf\oplus \gf$ (see \cite{LyakhovskyPostnova2011}). So our implementation of recurrent algorithm can be used to decompose tensor products (See Section \ref{sec:tens-prod-decomp}).

In the case of affine Lie algebras $\gf, \af$  the multiplicities
$m_{\nu}$ and the corresponding branching coefficients $b_{\nu}$
can be regarded as the coefficients in power series decomposition
of string and branching functions correspondingly:
\begin{align}
  \label{eq:21}
  &\sigma_{\nu}(q)=\sum_{n=0}^{\infty} m_{\nu-n\delta} q^n, \quad \nu=\sum_j c_j \omega_j,\quad c_j\geq 0\\
  & b_{\nu}(q)=\sum_{n=0}^{\infty} b_{\nu-n\delta} q^n,\quad  \nu=\sum_j c_j \omega_j, \quad c_j\geq 0
\end{align}
String and branching functions have  modular and analytic properties which are important for conformal field theory, especially in  coset models and the study of CFT on higher genus surfaces \cite{kac1988modular}, \cite{difrancesco1997cft}, \cite{Walton:1999xc}, \cite{walton1989conformal}.

{\bf Affine.m}  calculates weight multiplicities and branching coefficients for affine Lie algebras up to some finite grade. We present examples of computations in Sections \ref{sec:string-funct-affine}, \ref{sec:branch-funct-coset}. Now we proceed to the description of datastructures and algorithms implemented in {\bf Affine.m}.

 \section{Обозначения}

 \label{sec:notation}
 Пусть $\frak{g}$ и $\frak{a}$ -- (аффинные) алгебры Ли, и существует вложение  $\frak{a}\hookrightarrow \frak{g}$ такое, что  $\frak{a}$ -- редуктивная подалгебра в $\frak{g}$ с согласованным корневым пространством: $\frak{%
 h}_{\frak{a}}^{\ast }\subset \frak{h}_{\frak{g}}^{\ast }$. Мы используем следующие обозначения:

 $\frak{g=n}^{-}+\frak{h}+\frak{n}^{+}$ --- разложение Картана;

 $r$ , $\left( r_{\frak{a}}\right) $ --- ранг алгебра $\frak{g}$ $%
 \left( \mathrm{\text{соотв., }\frak{a}}\right) $ ;

 $\Delta $ $\left( \Delta _{\frak{a}}\right) $--- корневая система; $\Delta
 ^{+} $ $\left( \mathrm{\text{соотв., }\Delta _{\frak{a}}^{+}}\right) $--- набор положительных корней (алгебр $\frak{g}$ и $\frak{a}$ соответственно);

 $\mathrm{mult}\left( \alpha \right) $ $\left( \mathrm{mult}_{\frak{a}}\left(
 \alpha \right) \right) $ --- кратность корня $\alpha$ в $%
 \Delta $ (соотв., в $\left( \Delta _{\frak{a}}\right) $);

 $S\quad \left( S_{\frak{a}}\right) $ --- множество простых корней (для 
 $\gf$ и $\af$ соответственно);

 $\alpha _{i}$ , $\left( \alpha _{\left( \frak{a}\right) j}\right) $ ---  $%
 i$-й (соотв., $j$-й) простой корень алгебры $\frak{g}$ $\left( \mathrm{\text{соотв.,}\frak{a%
 }}\right) $; $i=0,\ldots ,r$,\ \ $\left( j=0,\ldots ,r_{\frak{a}}\right) $;

 $\alpha _{i}^{\vee }$ , $\left( \alpha _{\left( \frak{a}\right) j}^{\vee
 }\right) $-- простой ко-корень для $\frak{g}$ $\left( \mathrm{\text{соотв.,}\frak{a}%
 }\right) $ , $i=0,\ldots ,r$ ;\ \ $\left( j=0,\ldots ,r_{\frak{a}}\right) $;

 $W$ , $\left( W_{\frak{a}}\right) $--- группа Вейля;

 $C$ , $\left( C_{\frak{a}}\right) $--- фундаментальная камера Вейля;

 $\bar{C}, \left(\bar{C_{\frak{a}}}\right)$ --- замыкание фундаментальной камеры Вейля;

 $\epsilon \left( w\right) :=\left( -1\right) ^{\mathrm{length}(w)}$;

 $\rho $\ , $\left( \rho _{\frak{a}}\right) $\ --- вектор Вейля;

 $L^{\mu }$\ $\left( L_{\frak{a}}^{\nu }\right) $\ --- интегрируемый модуль  $\frak{g}$ со старшим весом $\mu $\ ; (соотв., интегрируемый $\af$-модуль старшего веса $\nu $);

 $\mathcal{N}^{\mu }$ , $\left( \mathcal{N}_{\frak{a}}^{\nu }\right) $ ---
 весовая диаграмма модуля $L^{\mu }$ (соотв., ${}L_{\frak{a}}^{\nu }$ );

 $P$ (соотв., $P_{\frak{a}} $) \ --- весовая решетка;

 $P^{+}$ (соотв., $P_{\frak{a}}^{+} $) \ --- решетка доминантных весов;

 $m_{\xi }^{\left( \mu \right) }$ , $\left( m_{\zeta }^{\left( \nu \right)
 }\right) $ --- кратность веса $\xi \in P$ \ $\left( \mathrm{\text{соотв., }\in P_{\frak{a}}}\right) $ в $L^{\mu }$, (соотв., в $\zeta \in
 L_{\frak{a}}^{\nu } $);

 $ch\left( L^{\mu }\right) $ (соотв., $\mathrm{ch}\left( L_{\frak{a}}^{\nu
 }\right) $)--- формальный характер $L^{\mu }$ (соответственно, $L_{\frak{a}}^{\nu
 } $);

 $ch\left( L^{\mu }\right) =\frac{\sum_{w\in W}\epsilon (w)e^{w\circ (\mu
 +\rho )-\rho }}{\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{%
 \mathrm{{mult}\left( \alpha \right) }}}$ --- формула Вейля-Каца;

 $R:=\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{{%
 mult}\left( \alpha \right) }}\quad $ (соотв., $R_{\frak{a}}:=\prod_{\alpha \in
 \Delta _{\frak{a}}^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{mult}_{\frak{a}%
 }\left( \alpha \right) } $)--- знаменатель Вейля.

\section{Конечномерные алгебры Ли}
\label{sec:finite-dimensional}
Основные определения: корни, веса, кратности.

\section{Теория представлений аффинных алгебр Ли}
\label{sec:representation-theory}

\subsection{Алгебра петель и центральное расширение}
\label{sec:loop-algebra-central-extension}

Определение аффинных алгебр. Корни, веса, кратности. Категория О.

\subsection{Представления аффинных алгебр Ли}
\label{sec:affine-lie-algebra-representations}

\section{Ветвления}
\label{sec:branching}
Веер вложения. Алгоритм.

\begin{abstract}
  Recurrent relations for branching coefficients in affine Lie algebras
  integrable highest weight modules are studied. The decomposition algorithm
  based on the injection fan technique is developed for the case of an arbitrary
  reductive subalgebra. In particular we consider the situation where
  the Weyl denominator becomes singular with respect to the subalgebra.
  We demonstrate
  that for any reductive subalgebra it is possible to define the
  injection fan and the analogue of the Weyl numerator -- the tools that describe
  explicitly the recurrent properties of branching coefficients.
  Possible applications of fan technique in CFT models are considered.
\end{abstract}

\subsection{Introduction}
\label{sec:introduction}

The branching problem for affine Lie algebras emerges in conformal field theory, for example,
in the construction of modular-invariant partition functions \cite{difrancesco1997cft}.
Recently the problem of conformal embeddings was considered in \cite{coquereaux2008conformal}.

There are different approaches to deal with branching coefficients. Some of them use the BGG
resolution \cite{bernstein1975differential} (for Kac-Moody algebras the algorithm is described in
\cite{kac1990idl},\cite{wakimoto2001idl}), the Schur function series \cite{fauser2006new}, the BRST
cohomology \cite{Hwang:1994yr}, Kac-Peterson formulas \cite{kac1990idl,quella2002branching} or some
combinatorial methods applied in \cite{feigin707principal}.

In this paper we prove that
for an arbitrary reductive subalgebra branching coefficients are subject to
a set of recurrent properties that can be explicitly formulated and that there exists
an effective and simple algorithm to solve these recurrent relations step by step.
The basic idea is similar to the one used in \cite{ilyin812pbc} for maximal embeddings.
In our case the algorithm is essentially different, new properties of singular weights
are determined to deal with an arbitrary reductive injection $\frak{a} \rightarrow \frak{g}$.

The principal point is to consider the subalgebra $\af$ together with its
counterpart $\afb$ orthogonal to $\af$.
For any reductive algebra $\af$ the subalgebra $\afb \subset \frak{g} $ is regular and reductive.
For a highest weight module $L^{\left( \mu \right)}$ and orthogonal pair of subalgebras
$\left(  \af, \afb \right)$ we consider
the so called singular element $\Psi^{\left( \mu \right)}$ (the numerator
in the Weyl character formula
$ch\left( L^{\mu }\right) =\frac{\Psi ^{\left( \mu \right) }}{\Psi ^{\left( 0\right) }}$,
see for example \cite{humphreys1997introduction})
the Weyl denominator $\Psi ^{\left( 0\right) }_{\afb}$ and the projection
$\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}
=\pi_{\af}\frac{\Psi ^{\left( \mu \right) }_{\frak{g}}}{\Psi ^{\left( 0\right) }_{\afb}}$.
We prove that for any highest weight $\hf$-diagonalizable module $L^{\left( \mu \right)}$ and orthogonal pair
$\left(  \af, \afb \right)$ the element
$\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}$ has a decomposition with respect to
the set of Weyl numerators $\Psi ^{\left( \mu \right) }_{ \afb }$ of $\afb$.
This decomposition provides the possibility to construct a recurrent property for branching coefficients corresponding
to the injection $\frak{a} \rightarrow \frak{g} $.
The property is formulated in
terms of the specific element $\Gamma_{\af \rightarrow \gf}$ of the group algebra
$\mathcal{E}\left( \frak{g} \right)$ called "the injection fan".
Using this tool we formulate a simple and
explicit algorithm for branching coefficients computations applicable for an arbitrary (maximal or nonmaximal)
subalgebras of finite-dimensional or affine Lie algebras.
In the case of maximal embedding the corresponding fan is unsubtracted, the singular element
becomes trivial
$\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}=\Psi ^{\left( \mu \right) }_{\left(  \gf\right)}$
and the relations described earlier in \cite{ilyin812pbc} are reobtained.

We demonstrate that our algorithm is effective and can be used in studies
of conformal embeddings and coset constructions in rational conformal field theory.

The paper is organized as follows. In the subsection \ref{sec:notation}  we fix the general notations.
In the Section \ref{sec:recurr-form-branch} we derive the decomposition formula based on
recurrent properties of anomalous branching coefficients and describe the decomposition algorithm
for integrable highest weight modules
$L_{\mathfrak{g}}$ with respect to a reductive subalgebra $\mathfrak{a}\subset \mathfrak{g}$
(subsection \ref{sec:algorithm}). In Section \ref{sec:finite-dimens-lie} we present several
simple examples for finite-dimensional Lie algebras. Affine Lie algebras and their applications in
CFT models are considered in Section \ref{sec:phys-appl}.
General properties of the proposed algorithm and
possible further developments are also discussed (Section \ref{sec:conclusion}).

\subsubsection{Notation}
\label{sec:notation}

Consider affine Lie algebras $\frak{g}$ and $\af$ with
underlying finite-dimensional subalgebras $\go$ and $%
\ao$ and an injection $\af\longrightarrow \frak{g%
}$ such that $\af$ is a reductive subalgebra $\frak{a\subset g}$ with
correlated root spaces: $\frak{h}_{\af}^{\ast }\subset \frak{h}_{\frak{g%
}}^{\ast }$ and $\frak{h}_{\ao}^{\ast }\subset \frak{h%
}_{\go}^{\ast }$\
.
We use the following notations:

$L^{\mu }$\ $\left( L_{\af}^{\nu }\right) $\ --- the integrable module
of $\frak{g}$ with the highest weight $\mu $\ ; (resp. integrable $\af$
-module with the highest weight $\nu $ );

$r$ , $\left( r_{\af}\right) $ --- the rank of the algebra $\frak{g}$ $%
\left( \mbox{resp. }\af\right) $ ;

$\Delta $ $\left( \Delta _{\af}\right) $--- the root system; $\Delta
^{+} $ $\left( \mbox{resp. }\Delta _{\af}^{+}\right) $--- the positive
root system (of $\frak{g}$ and $\af$ respectively);

$\mathrm{mult}\left( \alpha \right) $ $\left( \mathrm{mult}_{\af}\left(
\alpha \right) \right) $ --- the multiplicity of the root $\alpha$ in $\Delta
$ (resp. in $\left( \Delta _{\af}\right) $);

$\co{\Delta}$ , $\left( \co{\Delta _{\af}}%
\right)$ --- the finite root system of the subalgebra $\co{%
\frak{g}}$ (resp. $\co{\af}$);

$\mathcal{N}^{\mu }$ , $\left( \mathcal{N}_{\af}^{\nu }\right) $ --- the
weight diagram of $L^{\mu }$ $\left( \mbox{resp. }L_{\af}^{\nu }\right)
$ ;

$W$ , $\left( W_{\af}\right) $--- the corresponding Weyl group;

$C$ , $\left( C_{\af}\right) $--- the fundamental Weyl chamber;

$\bar{C}, \left(\bar{C_{\mathfrak{a}}}\right)$ --- the closure of the fundamental Weyl chamber;

$\rho $\ , $\left( \rho _{\af}\right) $\ --- the Weyl vector;

$\epsilon \left( w\right) :=\det \left( w\right) $ ;

$\alpha _{i}$ , $\left( \beta _{j}\right) $ --- the $i
$-th (resp. $j$-th) basic root for $\frak{g}$ $\left( \mbox{resp. }\af%
\right) $; $i=0,\ldots ,r$,\ \ $\left( j=0,\ldots ,r_{\af}\right) $;

$\delta $ --- the imaginary root of $\frak{g}$ (and of $\af$ if any);

$\alpha _{i}^{\vee }$ , $\left( \beta _{j}^{\vee
}\right) $--- the basic coroot for $\frak{g}$ $\left( \mbox{resp. }\af%
\right) $ , $i=0,\ldots ,r$ ;\ \ $\left( j=0,\ldots ,r_{\af}\right) $;

$\co{\xi }$ , $\co{\xi _{\left( \af\right) }}$
--- the finite (classical) part of the weight $\xi \in P$ , $\left( \mbox{%
resp. }\xi _{\left( \af\right) }\in P_{\af}\right) $;

$\lambda =\left( \co{\lambda };k;n\right) $ ---
decomposition of the affine weight $\lambda$ indicating the finite
part $\co{\lambda }$, the level $k$ and the grade $n$;

$P$ $\left( \mbox{resp. } P_{\af}\right) $ \ --- the weight lattice;

>>>>

\chapter{Рекуррентные соотношения для коэффициентов ветвления представлений аффинных алгебр Ли}
\label{cha:affine-lie-algebras}
Аффинные алгебры Ли - это естественное обобщение простых конечномерных алгебр Ли. Они возникают при изучении различных физических моделей, например, моделей Весса-Зумино-Новикова-Виттена в конформной теории поля и интегрируемых спиновых цепочек.

\section{Введение}
\label{sec:introduction}

Как было показано в разделе \ref{sec:modular-invariance}, проблема ветвления для аффинных алгебр Ли возникает в конформной теории поля, например, при построении модулярно-инвариантных статсумм  \cite{difrancesco1997cft}. Недавно проблема конформных вложений рассматривалась в работе \cite{coquereaux2008conformal}.

Существуют различные подходы к вычислению коэффициентов ветвления. Некоторые из них используют резольвенту Бернштейна-Гельфанда-Гельфанда \cite{bernstein1975differential} (См. главу \ref{cha:BGG}, алгоритм вычислений описан в работах \cite{kac1990idl},\cite{wakimoto2001idl}), ряды функций Шура \cite{fauser2006new}, когомологии БРСТ  \cite{Hwang:1994yr}, формулы Каца-Петерсона  \cite{kac1990idl,quella2002branching} или комбинаторные методы \cite{feigin707principal}.

В этой главе мы доказываем, что для произвольной редуктивной подалгебры коэффициенты ветвления описываются набором рекуррентных соотношений и существует эффективный и простой алгоритм для пошагового решения этих соотношений. Общая идея похожа на подход, предложенный в работе  \cite{ilyin812pbc} для максимальных вложений. Но в данном случае алгоритм существенно отличается, так как использует новые свойства сингулярных весов для работы с произвольными редуктивными вложениями $\frak{a} \rightarrow \frak{g}$.

При этом важно рассматривать подалгебру  $\af$ вместе с ее ``ортогональным партнером'' $\afb\subset \gf$. 
Для любой редуктивной подалгебры $\af$ подалгебра $\afb$ регулярна и редуктивна.

Для модуля старшего веса  $L^{\left( \mu \right)}$ и ортогональной пары подалгебр $\left(  \af, \afb \right)$ мы рассматриваем так называемый сингулярный элемент  $\Psi^{\left( \mu \right)}$ (числитель в формуле Вейля для характеров
$ch\left( L^{\mu }\right) =\frac{\Psi ^{\left( \mu \right) }}{\Psi ^{\left( 0\right) }}$,
см., например, \cite{humphreys1997introduction}), 
знаменатель Вейля $\Psi ^{\left( 0\right) }_{\afb}$ и проекцию
$\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}
=\pi_{\af}\frac{\Psi ^{\left( \mu \right) }_{\frak{g}}}{\Psi ^{\left( 0\right) }_{\afb}}$.

Мы доказываем, что для произвольного $\hf$-диагонализующего модуля старшего веса $L^{\left( \mu \right)}$ и ортогональной пары $\left(  \af, \afb \right)$ элемент
$\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}$ допускает разложение на множество числителей Вейля $\Psi ^{\left( \mu \right) }_{ \afb }$ модулей подалгебры $\afb$.
Это разложение дает возможность построить рекуррентное соотношение для коэффициентов ветвления, соответствующих вложению $\frak{a} \rightarrow \frak{g} $. Рекуррентное соотношение формулируется с использованием особого элемента  $\Gamma_{\af \rightarrow \gf}$ групповой алгебры
$\mathcal{E}\left( \frak{g} \right)$, который мы называем ``веером вложения''. 
Применение этих инструментов позволяет нам сформулировать простой и явный алгоритм для вычисления коэффициентов ветвления, который может быть использован в случае произвольных (максимальных и не максимальных) подалгебр конечномерных и аффинных алгебр Ли. 
В случае максимального вложения веер имеет простую форму, сингулярный элемент становится тривиальном $\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}=\Psi ^{\left( \mu \right) }_{\left(  \gf\right)}$ и мы восстанавливаем соотношения, полученные ранее в работе \cite{ilyin812pbc}.

Также мы показываем, что предложенный алгоритм эффективен и может использоваться при изучении конформных вложений и coset-моделей в рациональной конформной теории поля.

В следующем разделе \ref{sec:intro-simple-lie-algebras} мы определяем аффинные алгебры Ли и приводим основные определения в теории представлений.  Заодно мы фиксируем обозначения. Затем в разделе \ref{sec:branching} мы выводим формулы разложения, основывающиеся на аномальных коэффициентах ветвления и описываем алгоритм редукции интегрируемых модулей старшего веса $L_{\mathfrak{g}}$ по отношению к редуктивной подалгебре  $\mathfrak{a}\subset \mathfrak{g}$ (см. секцию \ref{sec:algorithm}). В разделе \ref{sec:branching-examples} мы представляем различные примеры применения алгоритма для конечномерных \ref{sec:finite-dimens-lie} и аффинных алгебр Ли и обсуждаем их роль в моделях конформной теории поля \ref{sec:phys-appl}. Выводы к главе изложены в разделе \ref{sec:conclusion}. В следующей главе \ref{cha:computational-methods} описывается реализация алгоритма на языке {\it Mathematica} и приводятся дополнительные примеры вычислений.

%%  \subsubsection{Notation}
%%  \label{sec:notation}
%%  
%%  Consider affine Lie algebras $\frak{g}$ and $\af$ with
%%  underlying finite-dimensional subalgebras $\go$ and $%
%%  \ao$ and an injection $\af\longrightarrow \frak{g%
%%  }$ such that $\af$ is a reductive subalgebra $\frak{a\subset g}$ with
%%  correlated root spaces: $\frak{h}_{\af}^{\ast }\subset \frak{h}_{\frak{g%
%%  }}^{\ast }$ and $\frak{h}_{\ao}^{\ast }\subset \frak{h%
%%  }_{\go}^{\ast }$\
%%  .
%%  We use the following notations:
%%  
%%  $L^{\mu }$\ $\left( L_{\af}^{\nu }\right) $\ --- the integrable module
%%  of $\frak{g}$ with the highest weight $\mu $\ ; (resp. integrable $\af$
%%  -module with the highest weight $\nu $ );
%%  
%%  $r$ , $\left( r_{\af}\right) $ --- the rank of the algebra $\frak{g}$ $%
%%  \left( \mbox{resp. }\af\right) $ ;
%%  
%%  $\Delta $ $\left( \Delta _{\af}\right) $--- the root system; $\Delta
%%  ^{+} $ $\left( \mbox{resp. }\Delta _{\af}^{+}\right) $--- the positive
%%  root system (of $\frak{g}$ and $\af$ respectively);
%%  
%%  $\mathrm{mult}\left( \alpha \right) $ $\left( \mathrm{mult}_{\af}\left(
%%  \alpha \right) \right) $ --- the multiplicity of the root $\alpha$ in $\Delta
%%  $ (resp. in $\left( \Delta _{\af}\right) $);
%%  
%%  $\co{\Delta}$ , $\left( \co{\Delta _{\af}}%
%%  \right)$ --- the finite root system of the subalgebra $\co{%
%%  \frak{g}}$ (resp. $\co{\af}$);
%%  
%%  $\mathcal{N}^{\mu }$ , $\left( \mathcal{N}_{\af}^{\nu }\right) $ --- the
%%  weight diagram of $L^{\mu }$ $\left( \mbox{resp. }L_{\af}^{\nu }\right)
%%  $ ;
%%  
%%  $W$ , $\left( W_{\af}\right) $--- the corresponding Weyl group;
%%  
%%  $C$ , $\left( C_{\af}\right) $--- the fundamental Weyl chamber;
%%  
%%  $\bar{C}, \left(\bar{C_{\mathfrak{a}}}\right)$ --- the closure of the fundamental Weyl chamber;
%%  
%%  $\rho $\ , $\left( \rho _{\af}\right) $\ --- the Weyl vector;
%%  
%%  $\epsilon \left( w\right) :=\det \left( w\right) $ ;
%%  
%%  $\alpha _{i}$ , $\left( \beta _{j}\right) $ --- the $i
%%  $-th (resp. $j$-th) basic root for $\frak{g}$ $\left( \mbox{resp. }\af%
%%  \right) $; $i=0,\ldots ,r$,\ \ $\left( j=0,\ldots ,r_{\af}\right) $;
%%  
%%  $\delta $ --- the imaginary root of $\frak{g}$ (and of $\af$ if any);
%%  
%%  $\alpha _{i}^{\vee }$ , $\left( \beta _{j}^{\vee
%%  }\right) $--- the basic coroot for $\frak{g}$ $\left( \mbox{resp. }\af%
%%  \right) $ , $i=0,\ldots ,r$ ;\ \ $\left( j=0,\ldots ,r_{\af}\right) $;
%%  
%%  $\co{\xi }$ , $\co{\xi _{\left( \af\right) }}$
%%  --- the finite (classical) part of the weight $\xi \in P$ , $\left( \mbox{%
%%  resp. }\xi _{\left( \af\right) }\in P_{\af}\right) $;
%%  
%%  $\lambda =\left( \co{\lambda };k;n\right) $ ---
%%  decomposition of the affine weight $\lambda$ indicating the finite
%%  part $\co{\lambda }$, the level $k$ and the grade $n$;
%%  
%%  $P$ $\left( \mbox{resp. } P_{\af}\right) $ \ --- the weight lattice;
%%  
%%  $m_{\xi }^{\left( \mu \right) }$ , $\left( m_{\xi }^{\left( \nu \right)
%%  }\right) $ --- the multiplicity of the weight $\xi \in P$ \ $\left( \mbox{%
%%  resp. }\in P_{\af}\right) $ in the module $L^{\mu }$ , (resp. $\xi \in
%%  L_{\af}^{\nu } $);
%%  
%%  $ch\left( L^{\mu }\right) $ $\left( \mbox{resp. }ch\left( L_{\af}^{\nu
%%  }\right) \right) $--- the formal character of $L^{\mu }$ $\left( \mbox{resp. }%
%%  L_{\af}^{\nu }\right) $;
%%  
%%  $ch\left( L^{\mu }\right) =\frac{\sum_{w\in W}\epsilon (w)e^{w\circ (\mu
%%  +\rho )-\rho }}{\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{%
%%  \mathrm{{mult}\left( \alpha \right) }}}$ --- the Weyl-Kac formula;
%%  
%%  $R:=\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{{%
%%  mult}\left( \alpha \right) }}\quad $
%%  $\left( \mbox{resp. }R_{\af}:=\prod_{\alpha \in \Delta _{%
%%  \af}^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{mult}_{\af}\mathrm{%
%%  \left( \alpha \right) }}\right) $--- the Weyl denominator.
%%  
%%  
%%  
\section{Простые алгебры Ли, алгебра петель, центральные расширения и аффинные алгебры Ли}

@@

\section{Рекуррентные соотношения для коэффициентов ветвления}
\label{sec:branching}
